---
title: "Physio Analysis"
format: html
editor: visual
bibliography: references.bib
---

## Libraries

```{r}
library(tidyverse)
library(here)
library(janitor)
library(googlesheets4)
```

## Load functions

```{r}
source(here("physio/scripts/functions.R"))

update_sample_id <- function(data){
  
  # make_long
  
  make_long <- function(sample_id){
    paste0(substr(sample_id, 1, 5),
           "0",
           substr(sample_id, 6, 8))
  }
  
  data_new <- data %>% 
    mutate(sample_id = ifelse(!is.na(sample_id) & nchar(sample_id) == 8, 
                              make_long(sample_id),
                              sample_id)
           )
  
  problems <- which(!is.na(data_new$sample_id) & nchar(data_new$sample_id) != 9)
  
  if (length(problems) > 0){
    warning(paste("Some sample_ids have not 9 digits:"), data_new$sample_id[problems])
  }
  
  return(data_new)
}



```

## Overview

```{r}
url_to_googlesheet <- "https://docs.google.com/spreadsheets/d/1Mli0yIHJowb9XL-YmBC_nRSJmSiv0pGbADODRgEGxBE/edit#gid=0"

dat_ov <- read_sheet(url_to_googlesheet, sheet = "1. Overview&24h")[,1:7] %>%
  clean_names() %>% 
  update_sample_id()
```

Check

```{r}
dat_ov %>% 
  filter(!is.na(sample_id) & duplicated(sample_id))
```

## Area

Get data

```{r}
dat_area <- read_sheet(url_to_googlesheet,
                       sheet = "4. Area") %>% 
  clean_names() %>% 
  update_sample_id()
```

Calculate Area by the increase in weigh between the first and second dip according to [@veal2010].

$$
A = 34.32 * \Delta w
$$

, where $A$ is the surface area in $cm^2$ and $\Delta w$ the the increase in weight in g.

```{r}
dat_area <- dat_area %>% 
  get_area()
```

Some checks

```{r}
dat_area %>% 
  filter(!is.na(sample_id) & duplicated(sample_id))
```

```{r}
dat_area %>% 
  pull(area) %>% 
  range(na.rm = T)
```

```{r}
dat_area %>% 
  pull(area) %>% 
  hist(breaks = 25, main = "Surface area in cm^2")
```

## Symbiodiniaceae Counts

```{r}
dat_counts <- read_sheet(url_to_googlesheet,
                         sheet = "3. Counts") %>% 
  clean_names() %>% 
  update_sample_id()
```

Calculate the counts for each replicate (6 replicate counts per sample).

The function calculates the overall counts per sample using the volume of the tissue slurry, the dilution before counting (1:3, `V_sw_added` and `V_sw_pipetted`), a second dilution step depending on the concentration (`V_zoox` and `V_sw`), and the volume used for counting (0.1 µL). It is then corrected for the surface area of the coral.

```{r}
dat_counts <- dat_counts %>% 
  get_counts(data_overview = dat_ov, 
             data_area = dat_area)
```

Checks

```{r}
dat_counts %>% 
  group_by(sample_id) %>% 
  summarise(n = n()) %>% 
  filter(n != 6)
```

```{r}
dat_counts %>% 
  pull(count_per_cm2) %>% 
  range(na.rm = T)
```

```{r}
dat_counts %>% 
  pull(count_per_cm2) %>% 
  hist(breaks = 25, main = "Symbiodiniaceae per cm^2")
```

## Chlorophyll a

Get overview data

```{r}
dat_chl24h_ov <- read_sheet(url_to_googlesheet,
                            sheet = "1. Overview&24h") %>% 
  clean_names() %>% 
  update_sample_id()

dat_chl48h_ov <- read_sheet(url_to_googlesheet,
                            sheet = "2. 48 h") %>% 
  clean_names() %>% 
  update_sample_id()
```

Read raw data from csv files and combine to one `data.frame` . Divide into data set after 24 h incubation and 48 h incubation time

```{r}
dat_chl <- read_all_files(here('physio/data/raw_data/during_bleaching'))

dat_chl24h <-  dat_chl %>% 
  filter(str_detect(filename, pattern = "24h"))

dat_chl48h <-  dat_chl %>% 
  filter(str_detect(filename, pattern = "48h"))


nrow(dat_chl24h) + nrow(dat_chl48h) == nrow(dat_chl)# ok
```

Calculate chlorophyll concentration ($µg~mL^{-1}$) for both incubation times based on the formulas given in [@jeffrey1975]:

$$
\text{Chlorophyll}~a = 11.43 \frac{A_{663} - A_{750}}{PL}  – 0.64\frac{A_{630} - A_{750}}{PL}
$$

, where $A_{663}$, $A_{630}$, and $A_{750}$ is the absorption at 663, 630, and 750 nm, and $PL$ the path length of the cuvette in cm.

The absorption at 750 nm is used as internal standard to correct for turbidity.

-   For each batch, the mean of the blanks are subtracted for the absorption.

-   The functions prints a summary table to assess if the blank values make sense. Check that they are not too high compared to the values of the samples

-   A $PL$ of 1 cm is assumed

-   The chlorophyll concentration is normalized for the area and has the unit $µg~cm^2$

### 24 h data

```{r}
dat_chl24h <- dat_chl24h %>% 
  get_chlorophyll(data_chlorophyll_overview = dat_chl24h_ov,
                  data_overview = dat_ov,
                  data_area = dat_area)
```

Checks

```{r}
dat_chl24h %>% 
  group_by(sample_id) %>% 
  summarise(n = n()) %>% 
  filter(n != 2)
```

```{r}
dat_chl24h %>% 
  pull(chl_a_per_cm2) %>% 
  range()
```

```{r}
dat_chl24h %>% 
  pull(chl_a_per_cm2) %>% 
  hist(breaks = 25, main = "Chlorophyll concentration")
```

### 48 h data

```{r}
dat_chl48h <- dat_chl48h %>% 
  get_chlorophyll(data_chlorophyll_overview = dat_chl48h_ov,
                  data_overview = dat_ov,
                  data_area = dat_area)
```

Checks

```{r}
dat_chl48h %>% 
  group_by(sample_id) %>% 
  summarise(n = n()) %>% 
  filter(n != 2)
```

```{r}
dat_chl48h %>% 
  pull(chl_a_per_cm2) %>% 
  range()
```

```{r}
dat_chl48h %>% 
  pull(chl_a_per_cm2) %>% 
  hist(breaks = 25, main = "Chlorophyll concentration")
```

### Combine 24h and 48 h

Compare

```{r}
dat_chl <- bind_rows(dat_chl24h %>% 
                       mutate(extr_time = "24 h"),
                     dat_chl48h %>% 
                       mutate(extr_time = "48 h"))

dat_chl %>% 
  ggplot(aes(x = sample_id, y = chl_a_per_cm2))+
  geom_point(aes(col = extr_time))+
  theme_minimal()


```

Add 24 h and 48 h measurements. The means need to be added (variability will be lost), but not possible in another way because m1 of 24 h does not correspond to m1 48 h and vice versa.

```{r}
dat_chlS <- dat_chl %>% 
  group_by(sample_id, extr_time) %>% 
  summarise(chl_a_per_cm2 = mean(chl_a_per_cm2, na.rm = T)) %>% 
  ungroup() %>% 
  pivot_wider(names_from = extr_time, values_from = chl_a_per_cm2) %>% 
  clean_names() %>% 
  mutate(chl_a_per_cm2 = x24_h + x48_h) %>% 
  select(sample_id, chl_a_per_cm2)

```

## Combine everything

```{r}
dat_all <- full_join(dat_counts %>% 
                       select(-bag_number), #otherwise, 2x in combined data
                     dat_chlS, 
                     by = "sample_id")
```

More samples than needed were analysed (the bag with the samples before the actual sampling). Select only the "good" samples by comparing to the overview file

```{r}
dat_ov_all <- read.csv(here("overview/data/2024_05_15_sample_list_Charlotte.csv"))
```

Only select `sample_id`s in `dat_all` if they occur in `dat_ov_all`:

```{r}
dat_all <- dat_all %>% 
  filter(sample_id %in% dat_ov_all$sample_id)
```

All samples there?

```{r}
dat_ov_all$sample_id[which(!dat_ov_all$sample_id %in% dat_all$sample_id)]
dat_all$sample_id[which(!dat_all$sample_id %in% dat_ov_all$sample_id)]
```

Yes

Add metadata

```{r}
dat_all <- dat_all %>% 
  get_info()
```

Save

```{r}
write.csv(dat_all, file = here("physio/data/clean/during_bleaching.csv"))
```

Some easypeasy plots

```{r}
dat_all %>% 
  ggplot(aes(x = count_per_cm2, y = chl_a_per_cm2))+
  geom_smooth(aes(col = site), method = "lm", se = F)+
  geom_point(aes(col = site), shape = 1)+
  labs(x = "# Symbiodiniaceae", y = "Chlorophyll a concentration")+
  facet_wrap(~spec)+
  theme_minimal()
```

```{r}
dat_all %>% 
  ggplot(aes(x = count_per_cm2, y = chl_a_per_cm2))+
  geom_smooth(aes(col = spec), method = "lm", se = F)+
  geom_point(aes(col = spec), shape = 1)+
  labs(x = "# Symbiodiniaceae", y = "Chlorophyll a concentration")+
  facet_wrap(~site)+
  theme_minimal()
```

```{r}
dat_all %>% 
  ggplot(aes(x = count_per_cm2, y = chl_a_per_cm2))+
  geom_smooth(aes(col = spec), method = "lm", se = F)+
  geom_point(aes(col = spec), shape = 1)+
  labs(x = "# Symbiodiniaceae", y = "Chlorophyll a concentration")+
  theme_minimal()
```

More chlorophyll per symbiont in *Acropora* than in *Pocillopora*?

```{r}
dat_all %>% 
  ggplot(aes(x = spec, y = chl_a_per_cm2))+
  geom_point(aes(col = site), shape = 1, position = position_jitterdodge())+
  stat_summary(geom = "errorbar", fun.data = mean_se, position = "dodge", aes(col = site), linewidth = 1)+
  labs(x = NULL, y = "Chlorophyll a concentration")+
  theme_minimal()
```

Why are all *A. hya* from E2B \> \~2?

```{r}
dat_all %>% 
  ggplot(aes(x = spec, y = count_per_cm2))+
  geom_point(aes(col = site), shape = 1, position = position_jitterdodge())+
   stat_summary(geom = "errorbar", fun.data = mean_se, position = "dodge", aes(col = site), linewidth = 1)+
  labs(x = NULL, y = "# Symbiodiniaceae")+
  theme_minimal() 
```
